---
title: "Hands-on Exercise 1"
author: "Lin Xinyue"
---

# Getting Started

## Install and launch R packages

The code chunk below uses p_load()of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman:: p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("Exam_data.csv")
```

# Plotting a simple bar chart

```{r}

ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
    
```

# Using geom_histogram

```{r}

ggplot(data=exam_data,
       aes(x = MATHS)) +
  geom_histogram()+
   theme_light()

```

# Colouring

```{r}

ggplot(data=exam_data, aes(x= MATHS)) + geom_histogram(bins=20,
color="black",
fill="lightyellow")+
 theme_light()
```

# Combine geom objects

```{r}

ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5) +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="pink",        
               size=4)  +
   theme_light()

```

# Using geom_violin

```{r}
ggplot(data=exam_data, 
       aes(y = ENGLISH, 
           x= GENDER)) +
  geom_violin()

```

# Scatterplot with a best fit curve

```{r}

ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              linewidth=0.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+
   theme_light()
```

# Using facet_wrap

```{r}

ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS) +
    theme_light()
```

# Using facet_grid

```{r}

ggplot(data=exam_data, 
       aes(x= SCIENCE)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS) +
    theme_light()
```

# Using Theme

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_light() +
  theme_minimal()
```

\`\`\`
