---
title: "Take home exercise 01"
author: "Lin Xinyue"
date: "February 1, 2025"
date-modified: "last-modified"
---

1.  

    1.  Preparing Data

1.2 Installing and loading the packages

```{r}
pacman::p_load(tidyverse, treemap, ggridges,
               ggcorrplot, ggstatsplot, GGally,
               treemapify, readxl,SmartEDA, easystats, gtsummary, ggstatsplot,
               ggExtra, ggiraph) 
```

1.3 Data Preparation

```{r}
shipperformance <- read_csv("Ship_Performance_Dataset.csv")
```

1.4 check the data type

```{r}
str(shipperformance) 
```

---all the data type are correct

1.5 Cheek the missing value

```{r}

char_cols <- sapply(shipperformance, is.character)
num_cols <- sapply(shipperformance, is.numeric)

none_count <- colSums(shipperformance[, char_cols] == "None")

na_count <- colSums(is.na(shipperformance[, num_cols]))

missing_summary <- data.frame(
  Column = c(names(none_count), names(na_count)),
  Missing_Values = c(none_count, na_count),
  Percentage = c(none_count / nrow(shipperformance) * 100, na_count / nrow(shipperformance) * 100)
)

print(missing_summary)
```

---under 5%

1.5 Rename the column rename the "Date" colunm to shown as quarter

```{r}
shipperformance <- shipperformance %>%
  mutate(Date = as.Date(Date, format="%Y-%m-%d"),  
         Quarter = paste0("Q", quarter(Date)),  
         Year = year(Date),                     
         Quarter_Year = paste0(Year, "-", Quarter)) 

shipperformance <- shipperformance %>%
  select(-Date, -Year)

head(shipperformance)

```



1.6 Validate Data
```{r}
shipperformance %>%
  ExpData(type = 1)
```
```{r}
shipperformance %>%
  ExpData(type = 2)
```
---
no zero variance variable and % of variables


2.  EDA

2.1 Check for the outliers


```{r}
shipperformance %>%
ExpNumViz(
  target = NULL,
  nlim = 10,
  Page = c(2,2))  


```
---

Check Outliers
```{r}

boxplot_data <- shipperformance %>%
  select(Revenue_per_Voyage_USD, Operational_Cost_USD, Efficiency_nm_per_kWh, Operational_Cost_USD,Turnaround_Time_hours) %>%
  pivot_longer(cols = everything(), names_to = "Metric", values_to = "Value")

ggplot(boxplot_data, aes(x = Metric, y = Value, fill = Metric)) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, alpha=0.6) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


3. Visual Analysis

3.1 Quarterly Revenue Trend Analysis
```{r}
ggplot(shipperformance, aes(x = Quarter_Year, y = Revenue_per_Voyage_USD, group = Ship_Type, color = Ship_Type)) +
  geom_line(stat="summary", fun="mean", size=1) +
  geom_point(stat="summary", fun="mean", size=2) +
  labs(title = "Quarterly Revenue Trend by Ship Type",
       x = "Quarter-Year",
       y = "Average Revenue per Voyage (USD)") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold", size = 18, color = "darkblue"))
```

```{r}
avg_revenue <- mean(shipperformance$Revenue_per_Voyage_USD, na.rm = TRUE)  # Calculate global average

ggplot(shipperformance, aes(x = Quarter_Year, y = Revenue_per_Voyage_USD, 
                            group = Ship_Type, color = Ship_Type)) +
  geom_smooth(stat = "summary", fun = "mean", method = "loess", se = TRUE, size = 1.2, alpha = 0.3) +  
  geom_point(stat = "summary", fun = "mean", size = 3, alpha = 0.7) +  
  geom_line(stat = "summary", fun = "mean", size = 1.5) +  
  geom_hline(yintercept = avg_revenue, linetype = "dashed", color = "darkgrey", size = 1) +  # Reference Line
  facet_wrap(~ Weather_Condition, scales = "free_y") +  
  labs(title = "Quarterly Revenue Trends with Reference Line",
       
       x = "Quarter-Year",
       y = "Average Revenue per Voyage (USD)",
       color = "Ship Type") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold", size = 18, color = "darkblue"),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
```


3.2 Two-sample Test

H0:There is no significant difference in the average fuel efficiency (nm per kWh) among different engine types.

H1:At least one engine type has significantly different fuel efficiency compared to the others.

```{r}
ggbetweenstats(
  data = shipperformance, 
  x = Engine_Type, 
  y = Efficiency_nm_per_kWh,
  type = "p" 
 )
```

3.3 ANOVA Test

H0:There is no significant difference in Average_Load_Percentage across different Maintenance_Status categories.

H1:There is a significant difference in Average_Load_Percentage across different Maintenance_Status categories.

```{r}
ggbetweenstats(
  data = shipperformance,
  x = Maintenance_Status,   
  y = Average_Load_Percentage, 
  type = "p",  
  pairwise.display = "s", 
  bf.message = TRUE 
)
```


3.4 Significant Test of Association (Depedence)
```{r}
ggbarstats(
  data = shipperformance,  
  x = Ship_Type,           
  y = Route_Type,         
  title = "Association Between Ship Type and Route Type"
)
```

3.5 Regression Model
```{r}
model <- lm(Efficiency_nm_per_kWh ~ Maintenance_Status +           Average_Load_Percentage + Turnaround_Time_hours + Engine_Power_kW, 
             data = shipperformance)
            

check_normality(model)  

```

```{r}
summary(model)
```
```{r}
check_collinearity(model)
```
```{r}
#| fig-width:12
#| fig-height:8
check_model(model)
```

```{r}
tbl_regression (model1,
                intercept = TRUE)
```


```{r}
p_model <- parameters(model)
```

```{r}
plot(parameters(model))
```

```{r}
ggcoefstats(model,
            output = "plot")
```










