---
title: "Take home exercise 01"
author: "Lin Xinyue"
date: "February 1, 2025"
date-modified: "last-modified"
---

# 1.Introduction

```         
The transport industry plays an increasingly important role in globalization and has a significant impact on sustainable development. One major challenge is finding ways to adjust various factors to improve maritime transportation efficiency while maximizing profits.

The transport industry plays an increasingly important role in globalization and has a significant impact on sustainable development. One major challenge is finding ways to adjust various factors to improve maritime transportation efficiency while maximizing profits.This dataset is designed for maritime analytics, allowing researchers and industry professionals to explore clustering, predictive modeling, and optimization strategies. It includes key variables such as shipping route type, total operational cost per voyage, and maintenance status, providing insights into ship performance under different conditions.

By analyzing ship performance, this dataset helps identify important factors affecting efficiency and supports better fleet management decisions. Understanding these relationships can contribute to reducing costs, improving fuel efficiency, and making maritime operations more sustainable.
```

# 2.Prepare Data

## 2.1Installing and loading the packages

The code chunk below uses the p_load() function to install and launch the packages in RStudio

```{r}
pacman::p_load(tidyverse, 
              ggcorrplot, ggstatsplot, GGally,
              SmartEDA, easystats, gtsummary, ggstatsplot,
              ggExtra,parallelPlot) 
```

-   tidyverse : Core collection of R packages designed for data science

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   ggcorrplot : Offer options for matrix reordering and significance level display

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   GGally : Create complex plots and simplify exploratory data analysis

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   SmartEDA : Automates exploratory data analysis

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   easystats : To simplify statistical modeling and analysis

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   gtsummary : Creates professional summary tables

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   ggstatsplot : Integrate statistical details, facilitating enriched data exploration

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

-   ggExtra : For adding marginal plots to ggplot2

    <hr style="border:none; border-top:1px solid #ddd; margin:20px 0;">

## 2.2 Import Data

This study utilizes the Ship Performance Clustering Dataset, sourced from Kaggle. The dataset provided in CSV format, consists of 2,736 rows and 18 columns, encompassing both numerical and categorical variables.

The dataset is publicly available under the CC BY 4.0 License, permitting reuse and modification with appropriate attribution.

The Ship Performance Dataset is a synthetic yet realistic compilation of operational metrics and attributes for vessels operating in the Gulf of Guinea

```{r}
shipperformance <- read_csv("Ship_Performance_Dataset.csv")
```

## 2.3 Understand the data

```{r}
shipperformance %>%
  ExpData(type = 1)
```

-   “No. of zero variance variables (uniform)” indicates that there is no variance among the data

-   "%. of variables having complete cases" means there are no missing values at all in the dataset.

## 2.4 Rename Column

The "Date" column was removed and replaced with a combination of "Quarter" and "Year". This transformation allows for a more structured and interpretable representation of time, facilitating trend identification and seasonal pattern recognition within the dataset.

```{r}
shipperformance <- shipperformance %>%
  mutate(Date = as.Date(Date, format="%Y-%m-%d"),  
         Quarter = paste0("Q", quarter(Date)),  
         Year = year(Date),                     
         Quarter_Year = paste0(Year, "-", Quarter)) 

shipperformance <- shipperformance %>%
  select(-Date, -Year)

head(shipperformance)

```

```{r}
shipperformance %>%
  ExpData(type = 2)
```

The newly processed dataset is now ready for analysis.

# 3.EDA

## 3.1 Explore the distribution

```{r}
shipperformance %>%
ExpNumViz(
  target = NULL,
  nlim = 10,
  Page = c(2,2))  


```

-   The density plots provide insights of the distribution characteristics of each variable, helping to identify their shape, skewness, and kurtosis. potential patterns, outliers and deviations.

-   All variables have skewness values close to zero, indicating relatively symmetric distributions. Kurtosis values suggest the distributions are slightly flatter than a normal distribution, meaning data points are more spread out.

## 3.2 Check the outliers

```{r}

boxplot_data <- shipperformance %>%
  select(Revenue_per_Voyage_USD, Operational_Cost_USD, Efficiency_nm_per_kWh, Operational_Cost_USD,Turnaround_Time_hours) %>%
  pivot_longer(cols = everything(), names_to = "Metric", values_to = "Value")

ggplot(boxplot_data, aes(x = Metric, y = Value, fill = Metric)) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, alpha=0.6) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

There are not outliers in the boxplot, meaning all values are in a reasonable range.

# 4. Visual Analysis

## 4.1 Quarterly Revenue Trends by Ship Type and Weather Impact Analysis

```{r}
ggplot(shipperformance, aes(x = Quarter_Year, y = Revenue_per_Voyage_USD, group = Ship_Type, color = Ship_Type)) +
  geom_line(stat="summary", fun="mean", size=1) +
  geom_point(stat="summary", fun="mean", size=2) +
  labs(title = "Quarterly Revenue Trend by Ship Type",
       x = "Quarter-Year",
       y = "Average Revenue per Voyage (USD)") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold", size = 18, color = "darkblue"))
```

```{r}
avg_revenue <- mean(shipperformance$Revenue_per_Voyage_USD, na.rm = TRUE)  # Calculate global average

ggplot(shipperformance, aes(x = Quarter_Year, y = Revenue_per_Voyage_USD, 
                            group = Ship_Type, color = Ship_Type)) +
  geom_smooth(stat = "summary", fun = "mean", method = "loess", se = TRUE, size = 1.2, alpha = 0.3) +  
  geom_point(stat = "summary", fun = "mean", size = 3, alpha = 0.7) +  
  geom_line(stat = "summary", fun = "mean", size = 1.5) +  
  geom_hline(yintercept = avg_revenue, linetype = "dashed", color = "darkgrey", size = 1) +  # Reference Line
  facet_wrap(~ Weather_Condition, scales = "free_y") +  
  labs(title = "Quarterly Revenue Trends with Reference Line",
       
       x = "Quarter-Year",
       y = "Average Revenue per Voyage (USD)",
       color = "Ship Type") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold", size = 18, color = "darkblue"),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
```

::: info-box
<h4>

</h4>

<ul>The first visualization shows quarterly revenue changes for different ship types (distinguished by color). Container ships maintained relatively stable revenue across quarters, while fish carriers and bulk carriers exhibited significant fluctuations, suggesting that certain ship types may be more sensitive to market conditions or operational efficiency. The second graph incorporates weather conditions. The calm & moderate weather correlates with stable revenue, but rough weather increases volatility, especially for fish and bulk carriers. This implies weather impacts operational efficiency. It should be consider for ship type allocation to reduce revenue fluctuations.</ul>
:::

## 4.2 Engine Type Efficiency Analysis

H0:There is no significant difference in the average fuel efficiency (nm per kWh) among different engine types.

H1:At least one engine type has significantly different fuel efficiency compared to the others.

```{r}
ggbetweenstats(
  data = shipperformance, 
  x = Engine_Type, 
  y = Efficiency_nm_per_kWh,
  type = "p" 
 )
```

::: info-box
<h4>

</h4>

<ul>Welch’s ANOVA (F = 3.92, p = 0.0087) shows significant efficiency differences (nm/kWh) among engines. Steam turbines are slightly better but not significant. The effect size (9.09e-03) suggests other factors matter which reminder alternative engines but also need to consider overall factors.</ul>
:::

## 4.3 Relationship Between Maintenance Status and Cargo Load

H0:There is no significant difference in Average_Load_Percentage across different Maintenance_Status categories.

H1:There is a significant difference in Average_Load_Percentage across different Maintenance_Status categories.

```{r}
ggbetweenstats(
  data = shipperformance,
  x = Maintenance_Status,   
  y = Average_Load_Percentage, 
  type = "p",  
  pairwise.display = "s", 
  bf.message = TRUE 
)
```

::: info-box
<h4>

</h4>

<ul>The ANOVA test is evaluate whether Average Load Percentage varies significantly across different Maintenance Status categories.The results (F = 1.29, p = 0.28) indicates that there is no significant differences in average load percentage across maintenance statuses. Therefore, improve maintenance alone may not enhance cargo load efficiency.</ul>
:::

## 4.4 Association Between Ship Type and Route Type

```{r}
ggbarstats(
  data = shipperformance,  
  x = Ship_Type,           
  y = Route_Type,         
  title = "Association Between Ship Type and Route Type"
)
```

::: info-box
<h4>

</h4>

<ul>A stacked bar chart showed consistent proportions of ship types (Bulk Carriers, Container Ships, etc.) across route types.p-value is 0.79 in Chi-square test indicates there is no significant relationship between Ship Type and Route Type. As the result, ship types are viable across all route types and the company has the flexibility to adjust the ship types according to different needs.</ul>
:::

## 4.5 Regression Analysis of Efficiency Determinants

```{r}
model <- lm(Efficiency_nm_per_kWh ~ Maintenance_Status +
              Average_Load_Percentage + 
              Turnaround_Time_hours + 
              Engine_Power_kW, 
             data = shipperformance)
            

check_normality(model)  

```

```{r}
summary(model)
```

```{r}
check_collinearity(model)
```

```{r}

check_model(model)
```

```{r}
tbl_regression (model,
                intercept = TRUE)
```

```{r}
p_model <- parameters(model)
```

```{r}
plot(parameters(model))
```

```{r}
ggcoefstats(model,
            output = "plot")
```

::: info-box
<h4>

</h4>

<ul>A regression model tested the impact of Maintenance Status, Average Load Percentage, Turnaround Time, and Engine Power on efficiency (nm/kWh). None of the variables showed significant effects (p \> 0.05) and the R-squared value is 0.00176 indicating poor model fit. For better predictions in the future, other factors for example operation cost may need to consider.</ul>
:::
