---
title: "Take-home Exercise 03"
author: "Lin Xinyue"
---

1.  Prepare the data

1.1 Install and load the packages

```{r}
pacman::p_load(tidyverse,dplyr,tidyr, readxl,writexl,ggplot2, ggcorrplot, ggstatsplot, GGally,
              SmartEDA, easystats, gtsummary,
              ggExtra,parallelPlot,ggdist, ggridges, ggthemes,
               colorspace,lubridate, ggthemes, reactable,
              reactablefmtr, gt, gtExtras)
```

1.2 Import the data

```{r}
water_indicators <- read.csv("WDI Country Data.csv")

file_path <- "Waterand Disease by WHO Region.xlsx"

```

```{r}
all_sheets <- setNames(
  lapply(excel_sheets(file_path), function(sheet) read_excel(file_path, sheet = sheet)),
  excel_sheets(file_path)
)
typhoid <- all_sheets[["typhoid 2000 - 2021"]]
diarrhea <- all_sheets[["diarrheal diseases 2000 - 2021"]]
hepatitiesA <- all_sheets[[ "acute hepatitis A 2000 -2021" ]]
riskfactor <- all_sheets[[ "Risk factor" ]]

```

1.3 Check the data

```{r}
water_indicators %>%
  ExpData(type = 2)
```

```{r}
typhoid %>%
  ExpData(type = 2)
```

```{r}
diarrhea %>%
  ExpData(type = 2)
```

```{r}
hepatitiesA %>%
  ExpData(type = 2)
```

```{r}
riskfactor %>%
  ExpData(type = 2)
```

1.4 remove the missing value

```{r}
water_indicators[water_indicators == ".."] <- NA
water_indicators[water_indicators == "#N/A"] <- NA
water_indicators[water_indicators == ""] <- NA

```

```{r}
water_indicators %>%
  ExpData(type = 2)
```

```{r}
water_indicators <- water_indicators[complete.cases(water_indicators[, 5:27]), ]
```

```{r}
riskfactor[riskfactor == " "] <- NA
riskfactor[riskfactor == "#N/A"] <- NA
riskfactor <- riskfactor[rowSums(is.na(riskfactor)) != ncol(riskfactor), ]
hepatitiesA[hepatitiesA == " "] <- NA
diarrhea[diarrhea == " "] <- NA
typhoid[typhoid == " "] <- NA
```

```{r}

riskfactor <- riskfactor[, 1:10]
hepatitiesA <- hepatitiesA[complete.cases(hepatitiesA), ]
diarrhea <- diarrhea[complete.cases(diarrhea), ]
typhoid <- typhoid[complete.cases(typhoid), ]

```

1.5 change the data type

```{r}
water_indicators[6:27] <- lapply(water_indicators[6:27], function(x) as.numeric(as.character(x)))
```

1.6 data summary

```{r}
water_indicators %>%
  ExpData(type = 2)
```

check how many countries in data

```{r}
length(unique(water_indicators$Country.Name))
```

1.7 reshape the data format

```{r}

water_indicators <- water_indicators %>%
  pivot_longer(
    cols = starts_with("X"),             
    names_to = "Year",
    values_to = "Value"
  )
```

```{r}
water_indicators <- water_indicators %>%
  mutate(Year = gsub("X(\\d{4})\\.\\..*", "\\1", Year),
         Year = as.numeric(Year),
         Value = as.numeric(Value))  
```

1.8 change the colunm name

```{r}
typhoid <- typhoid %>% rename(Region = Location)
diarrhea <- diarrhea %>% rename(Region = Location)
hepatitiesA <- hepatitiesA %>% rename(Region = Location)
riskfactor <- riskfactor %>% rename(Region = Location)
```

2.  Data Explore

```{r}

ggplot(typhoid, 
       aes(x = Value, 
           y = Region)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 0.01,  # Adjust depending on your rate scale
    fill = lighten("#7097BB", 0.3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "Typhoid Rate",
    expand = c(0, 0)
  ) +
  scale_y_discrete(
    name = NULL,
    expand = expansion(add = c(0.2, 2.6))
  ) +
  theme_ridges() +
  labs(title = "Distribution of Typhoid Rates by WHO Region")
```

```{r}
ggplot(diarrhea, 
       aes(x = Region, 
           y = Value)) +
  stat_halfeye(
    adjust = 0.9,              
    justification = -0.25,
    .width = 0,
    point_colour = NA,
    fill = "#7097BB"
  ) +
  geom_boxplot(
    width = 0.15,
    outlier.shape = NA,
    alpha = 0.6
  ) +
  stat_dots(
    side = "left",
    justification = 2.2,
    binwidth = 0.08,          
    dotsize = 2.2,
    color = "#7097BB",
    alpha = 0.9
  ) +
  coord_flip() +
  theme_economist() +
  labs(
    title = "Typhoid Rates by WHO Region"
  )

```

Group by Region 给region group by

```{r}
water_region <- water_indicators %>%
  group_by(Region, Year, Series.Name) %>%
  summarise(AvgValue = mean(Value, na.rm = TRUE), .groups = "drop")
```



```{r}
Typhoid <- typhoid %>%
   select(Region, Year, TyphoidRate = Value)

Diarrhea <- diarrhea %>%
   select(Region, Year, DiarrheaRate = Value)

HepatitisA <- hepatitiesA %>%
   select(Region, Year, HepatitisRate = Value)

Risk <- riskfactor %>%
   select(Region, Year, UnsafeRisk = Value)

```

```{r}
disease_combined <- reduce(
  list(Typhoid, Diarrhea, HepatitisA, Risk),
  ~ full_join(.x, .y, by = c("Region", "Year"))
)
```

**Prepare the whole data**

```{r}
merged_data <- left_join(disease_combined, water_region, by = c("Region", "Year"))
```

```{r}
merged_data <- merged_data[, !names(merged_data) %in% 
  "People with basic handwashing facilities including soap and water (% of population)"]

merged_data <- merged_data[merged_data$Region != "WHO region", ]
```

```{r}
merged_data <- merged_data %>%
  filter(!is.na(Year)) %>%
  mutate(Year = as.numeric(Year))
```

```{r}

ggplot(diarrhea,
       aes(x = Value, 
           y = Region, 
           fill = factor(stat(quantile)))) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = 6,
    quantile_lines = TRUE,
    scale = 2,
    alpha = 0.9
  ) +
  scale_fill_viridis_d(name = "Quartiles") +
  labs(
    title = "Typhoid Rate Distribution by WHO Region"
  ) +
  theme_ridges()
```

```{r}
filtered_corr_data <- merged_data %>%
  filter(`Series.Name` == "People using at least basic sanitation services (% of population)")

ggscatterstats(
  data = filtered_corr_data,
  x = AvgValue,        # water indicator
  y = TyphoidRate,            # disease rate
  xlab = "Basic Sanitation Access (%)",
  ylab = "Disease Rate",
  title = "Correlation between Basic Sanitation and Disease Rate",
  marginal = TRUE
)
```

***4. CDA-Correlation Analysis*** **corelationship analysis** *Drinking Water Access vs Typhoid Rate*

```{r}
df_plot1 <- merged_data %>%
  
  filter(Series.Name == "People using at least basic drinking water services (% of population)")
  "TyphoidRate"

ggscatterstats(
  data = df_plot1,
  x = AvgValue,
  y = TyphoidRate,
  marginal = TRUE,
  xlab = "Basic Drinking Water Access (%)",
  ylab = "Typhoid Rate",
  title = "Correlation: Drinking Water Access vs Typhoid Rate",
  point.args = list(alpha = 0.3)
)
```

Diarrhea Rate vs. Basic Water Access

```{r}
df_plot2 <- merged_data %>%
  filter(Series.Name == "People using at least basic drinking water services (% of population)")
   "DiarrheaRate"
   
  ggscatterstats(
    data = df_plot2,
    x = AvgValue,
    y = DiarrheaRate,
    title = "Basic Water Access vs. Diarrhea Rate",
    marginal = TRUE,
    point.args = list(alpha = 0.3)
  )
```

**ANOVA** *Basic Sanitation Levels Across Diseases*

```{r}
# Step 1: Pivot disease columns into long format
ANOVA1_data <- merged_data %>%
  tidyr::pivot_longer(
    cols = c(TyphoidRate, DiarrheaRate, HepatitisRate, UnsafeRisk),
    names_to = "Disease",
    values_to = "Rate"
  )
# Step 2: Filter for a specific water indicator
basic_sanitation_data <- ANOVA1_data  %>%
  filter(Series.Name == "People using at least basic sanitation services (% of population)")
# Step 3: Run ANOVA across disease groups
ggbetweenstats(
  data = basic_sanitation_data,
  x = Disease,
  y = AvgValue,
  type = "parametric",  # You can try "nonparametric" or "robust"
  title = "ANOVA: Basic Sanitation Levels Across Diseases",
  messages = FALSE
)
```

*Diarrhea Rate Across Sanitation Access Levels*

```{R}

# Step 1: Pivot to long format for diseases
long_data <- merged_data %>%
  tidyr::pivot_longer(
    cols = c(TyphoidRate, DiarrheaRate, HepatitisRate, UnsafeRisk),
    names_to = "Disease",
    values_to = "Rate"
  )

# Step 2: Filter for Diarrhea only and sanitation indicator
diarrhea_sanitation <- long_data %>%
  filter(Disease == "DiarrheaRate",
         `Series.Name` == "People using at least basic sanitation services (% of population)")

# Step 3: Create bins of sanitation level
diarrhea_sanitation <- diarrhea_sanitation %>%
  mutate(SanitationLevel = cut(
    AvgValue,
    breaks = c(0, 50, 70, 90, 100),
    labels = c("Low", "Moderate", "High", "Very High"),
    right = FALSE
  ))

# Step 4: Run ANOVA to compare Diarrhea Rate across sanitation groups
ggbetweenstats(
  data = diarrhea_sanitation,
  x = SanitationLevel,
  y = Rate,
  type = "parametric",
  title = "ANOVA: Diarrhea Rate Across Sanitation Access Levels"
)

```

*Time Trend* For water indicators

```{r}

library(dplyr)
library(gt)
library(gtExtras)

# Select indicator
selected_indicator <- "People using at least basic sanitation services (% of population)"

# Filter and summarize
water_summary_all_regions <- merged_data %>%
  filter(`Series.Name` == selected_indicator) %>%
  group_by(Region) %>%
  summarise(
    Min = min(AvgValue, na.rm = TRUE),
    Max = max(AvgValue, na.rm = TRUE),
    Average = round(mean(AvgValue, na.rm = TRUE), 1),
    Monthly = list(AvgValue[order(Year)]),
    Actual = mean(AvgValue[Year == 2021], na.rm = TRUE),
    Target = 100,
    .groups = "drop"
  )

# Display with sparkline + bullet chart
water_summary_all_regions %>%
  gt() %>%
  gt_plt_sparkline(Monthly, type = "default") %>%  # <-- FIXED HERE
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", "black")) %>%
  gt_theme_espn()


 
```

For disease

```{r}
library(dplyr)
library(gt)
library(gtExtras)

# Set the column name of the disease
selected_disease_column <- "DiarrheaRate"

# Filter only rows with non-missing disease values
disease_summary_all_regions <- merged_data %>%
  filter(!is.na(.data[[selected_disease_column]])) %>%
  group_by(Region) %>%
  summarise(
    Min = min(.data[[selected_disease_column]], na.rm = TRUE),
    Max = max(.data[[selected_disease_column]], na.rm = TRUE),
    Average = round(mean(.data[[selected_disease_column]], na.rm = TRUE), 1),
    Monthly = list(.data[[selected_disease_column]][order(Year)]),
    Actual = mean(.data[[selected_disease_column]][Year == 2021], na.rm = TRUE),
    Target = 0,
    .groups = "drop"
  )

# Generate the table
disease_summary_all_regions %>%
  gt() %>%
  gt_plt_sparkline(Monthly, type = "default") %>%
  gt_plt_bullet(column = Actual,
                target = Target,
                width = 28,
                palette = c("salmon", "black")) %>%
  gt_theme_espn()

```

```{r}
library(dplyr)
library(ggplot2)

# Example: Diarrhea Rate over time
trend_data <- merged_data %>%
  filter(!is.na(DiarrheaRate)) %>%
  group_by(Region, Year) %>%
  summarise(mean_disease = mean(DiarrheaRate, na.rm = TRUE), .groups = "drop")
```

```{r}
ggplot(trend_data, aes(x = Year, y = mean_disease, color = Region)) +
  geom_line(size = 1) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +
  labs(title = "Trend of Diarrhea Rate Over Time by Region",
       y = "Mean Diarrhea Rate (%)",
       x = "Year") +
  theme_minimal()
```

```{r}
ggplot(trend_data, aes(x = Year, y = mean_disease)) +
  geom_line(color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +
  facet_wrap(~ Region) +
  labs(title = "Time Trends by Region", y = "Disease Rate", x = "Year") +
  theme_minimal()
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# 1. Choose indicator + disease
selected_water_indicator <- "People using at least basic sanitation services (% of population)"
selected_disease <- "DiarrheaRate"

# 2. Prepare data
combined_trend_data <- merged_data %>%
  filter(Series.Name == selected_water_indicator) %>%
  select(Region, Year, WaterValue = AvgValue, !!selected_disease) %>%
  rename(DiseaseValue = !!selected_disease) %>%
  filter(!is.na(WaterValue) & !is.na(DiseaseValue))

# 3. Reshape for plotting
plot_data <- combined_trend_data %>%
  pivot_longer(cols = c(WaterValue, DiseaseValue),
               names_to = "Variable",
               values_to = "Value")

# 4. Plot
ggplot(plot_data, aes(x = Year, y = Value, color = Variable)) +
  geom_line(size = 1.2) +
  facet_wrap(~ Region) +
  labs(title = "Combined Trends: Water Access & Disease Burden Over Time",
       y = "Value (%)",
       x = "Year",
       color = "Metric") +
  theme_minimal()
```
