---
title: "Hands-on Exercise 2"
author: "Lin Xinyue"
---

# **2. Beyond ggplot2 Fundamentals**

## **2.1 Getting started**

Installing and loading the required libraries

```{r}
pacman:: p_load(tidyverse)
```

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

## 2.2 Importing data

```{r}
exam_data <- read_csv("Exam_data.csv", col_types = cols(
  ID = col_character(),
  CLASS = col_character(),
  GENDER = col_character(),
  RACE = col_character(),
  ENGLISH = col_double(),
  MATHS = col_double(),
  SCIENCE = col_double()
))
```

## 2.3 Beyand ggplot2 Annotation: ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### 2.3.1 working with ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## 2.4 Beyond ggplot2 Themes

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### 2.4.1 Working with ggthems package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### 2.4.2 Working with hrbthems package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_classic()
```

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_minimal(base_size = 15) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15),  
        panel.grid.major.y = element_line(),  
        panel.grid.minor.y = element_blank())

```

## 2.5 Beyond Single Graph

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

```

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### 2.5.1 About pathwork methods The general syntax that combines:

-   Two-Column Layout using the Plus Sign +

-   Parenthesis () to create a subplot group

-   Two-Row Layout using the Division Sign /

### 2.5.2 **Combining two ggplot2 graphs**

```{r}
p1 + p2
```

### 2.5.3 **Combining three ggplot2 graphs**

```{r}
(p1 / p2) | p3
```

### 2.5.4 **Creating a composite figure with tag**

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### 2.5.5 **Creating figure with insert**

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

### 2.5.6 **Creating a composite figure by using patchwork and ggtheme**

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
