---
title: "Hands-on Ex05_P4"
author: "Lin Xinyue"
---
1.  Preparing Data

1.2 Installing and loading the packages

```{R}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

1.3 Data Preparation
```{r}
wh <- read_csv("WHData-2018.csv") 
```

2.Plotting Static Parallel Coordinates Plot
To plot a basic static parallel coordinates plot by using ggparcoord().
```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

2.1 Plotting a parallel coordinates with boxplot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

2.2 Parallel coordinates with facet
```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

2.3 Rotating x-axis text label
```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```
2.4 Adjusting the rotated x-axis text label
```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

3. Plotting Interactive Parallel Coordinates Plot: parallelPlot methods
3.1 The basic plot

```{r}
wh <- wh %>%
  dplyr:::select.data.frame(., "Happiness score", 7:12)

parallelPlot(wh,
             width = 320,
             height = 250)
```

3.2 Rotate axis label
```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

3.3 Changing the colour scheme
```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

3.4 Parallel coordinates plot with histogram
```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

             






