---
title: "Hands-on Exercise 05_P2"
author: "Lin Xinyue"
---

1.  Preparing Data

1.2 Installing and loading the packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

1.3 Importing Data

```{r}
wine <- read_csv("wine_quality.csv")
```

2.  Building Correlation Matrix: pairs() method

```{r}
pairs(wine[,1:11])
```

```{r}
pairs(wine[,2:12])
```

2.1 Drawing the lower corner

```{r}
pairs(wine[,2:12], lower.panel = NULL)
```

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

2.2 Including with correlation coefficients

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

3 Visualising Correlation Matrix: ggcormat()

Corrgram data visualisation can be use when the number of observations is relastively large(more than 500)

ggcorrmat() of ggstatsplot package will be used.

3.1 The basic plot

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11)
```

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

-   cor.vars argument is used to compute the correlation matrix needed to build the corrgram.

-   ggcorrplot.args argument provide additional (mostly aesthetic) arguments that will be passed to ggcorrplot::ggcorrplot function. The list should avoid any of the following arguments since they are already internally being used: corr, method, p.mat, sig.level, ggtheme, colors, lab, pch, legend.title, digits.

For control the specific component of the plot such as the font size of the x-axis, y-axis, and the statistical report.

```{r}
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```

3.2 Building multiple plots

By using grouped_ggcorrmat() of ggstatsplot

```{r}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

::: info-box
```         
<h4>Note</h4>
<ul>
    <li>to build a facet plot, the only argument needed is grouping.var.</li>
    <li>Behind group_ggcorrmat(), patchwork package is used to create the multiplot. plotgrid.args argument provides a list of additional arguments passed to patchwork::wrap_plots, except for guides argument which is already separately specified earlier.
```

</li>

<li>

Likewise, annotation.args argument is calling plot annotation arguments of patchwork package.

<li>

</ul>

</div>

4.  Visualising Correlation Matrix by using corrplot Package 4.1 Getting started with corrplot

```{r}
wine.cor <- cor(wine[, 1:11])
```

corrplot() is used to plot the corrgram by using all the default setting as shown in the code chunk below.

```{r}
corrplot(wine.cor,tl.col = "black")
```

The default visual object used to plot the corrgram is circle and the default layout of the corrgram is a symmetric matrix.

4.2 Working with visual geometrics In corrplot package, there are seven visual geometrics can be used to encode the attribute values: circle, square, ellipse, number, shade, color and pie But can be changed, for example

```{r}
corrplot(wine.cor, tl.col = "black",method = 'shade', order = 'AOE', diag = FALSE)
```

4.3 Working with layout

corrplor() supports three layout types, namely: “full”, “upper” or “lower”. Foe example

```{r}
corrplot(wine.cor, tl.col = "black",
         method = "ellipse", 
         type="lower",
         tl.pos = "r",
         tl.srt = 0
         )
```

5.  Working with mixed layout

corrplot.mixed() can design corrgram with mixed visual matrix of one half and numerical matrix on the other half.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

6.Combining corrgram with the significant test

cor.mtest() to compute the p-values and confidence interval for each pair of variables

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

then use the p.mat argument of corrplot function

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         pch.col = "lightgreen",
         sig.level = .05)
```

7.  Reorder a corrgram

corrplot package support four sorting methods

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

7.1 Reordering a correlation matrix using hclust corrplot() can draw rectangles around the corrgram based on the results of hierarchical clustering with it.

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```

:::

:::
:::
