---
title: "Hands-on Exercise_P3"
author: "Lin Xinyue"
---
1.  Preparing Data

1.2 Installing and loading the packages

```{R}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

1.3 Data Preparation
```{r}
pop_data <- read_csv("WHData-2018.csv") 
```

1.4 Preparing the data
```{r}
wh <- read.csv("WHData-2018.csv", header = TRUE)

row.names(wh) <- wh$Country
```

2. Transforming the data frame into a matrix
Transfer data frame to data matrix to make your heatmap
```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

3. Static Heatmap 
3.1 heatmap() of R Stats
```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```


```{r}
wh_heatmap <- heatmap(wh_matrix)
```
```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```
4. Creating Interactive Heatmap
```{r}
heatmaply(mtcars)
```

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

5. Data trasformation
Three main data transformation methods are supported by heatmaply(), namely: 
 scale, normalise and percentilse.

5.1 Scaling method
It would bring them all close to the standard normal distribution.
The scale argument in heatmaply() supports column and row scaling.
```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

5.2 Normalising method

It can be used to bring data to the 0 to 1 scale by subtracting the minimum and dividing by the maximum of all observations.
```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```
5.3 Percentising method

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

6. Clustering algorithm
6.1 Manual approach
```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```
6.2 Statistical approach
dend_expend() will be used to determine the recommended clustering method to be used.
```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

find_k() is used to determine the optimal number of cluster.
```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

7. Seriation

Optimal Leaf Ordering (OLO)
```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

Gruvaeus and Wainer(GW)

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```


The option “mean” gives the output we would get by default from heatmap functions in other packages such as gplots::heatmap.2.
```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

“none” gives us the dendrograms without any rotation that is based on the data matrix.
```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

8. Working with colour palettes

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

9. Interactive heatmap

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```

































