---
title: "Take-Home Exercise 02"
author: "Lin Xinyue"
date: "March 03, 2025"
date-modified: "last-modified"
---
1.Data Preparation 

1.1 Installing R pacakges
```{r}
pacman::p_load(tidyverse,readxl, gifski, gapminder,
               plotly, gganimate, lubridate,
               scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               CGPfunctions, ggHoriPlot, gtsummary,
               easystats,tidyverse, GGally,SmartEDA,
               treemapify,treemap,d3treeR,
               ggExtra,parallelPlot)
```

1.2 Importing the data

```{r}
merchandisetrade <- read_xlsx("Total_Merchandise_Trade_201525.xlsx")
```

手动去掉了文字解释后

1.3 Examining the data structure

```{r}
kable(head(merchandisetrade))
```


1.4 Data Preparation

For merchandise 

```{r}
month_levels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

Year_merchandise <- merchandisetrade %>%
  pivot_longer(
    cols = matches("\\d{4}\\s[A-Za-z]{3}"), 
    names_to = "date_col",
    values_to = "value"
  ) %>%
  mutate(
    year = as.numeric(str_extract(date_col, "\\d{4}")),
    month = str_extract(date_col, "[A-Za-z]{3}")
  ) %>%
  select(-date_col) %>%
  arrange(year, match(month, month_levels)) %>%
  mutate(month = factor(month, levels = month_levels))

```

```{r}
Year_merchandise <- Year_merchandise %>%
  group_by(`Data Series`, year) %>% 
  summarize(yearly_total = sum(value, na.rm = TRUE), .groups = "drop")
```

```{r}
kable(head(Year_merchandise))
```
```{r}
Year_merchandise %>%
  ExpData(type = 2)
```

visulization

```{r}
Year_merchandise <- Year_merchandise %>%
  filter(year >= 2018 & year <= 2024, 
         grepl("Total Merchandise (Trade|Exports|Imports)", `Data Series`))

Year_merchandise <- Year_merchandise %>%
  mutate(Category = case_when(
    grepl("Total Merchandise Trade", `Data Series`) ~ "Total Trade",
    grepl("Total Merchandise Exports", `Data Series`) ~ "Exports",
    grepl("Total Merchandise Imports", `Data Series`) ~ "Imports"
  )) %>%
  select(year, yearly_total, Category, `Data Series`)

trade_growth <- Year_merchandise %>%
  filter(Category == "Total Trade") %>%
  arrange(year) %>%
  mutate(growth_pct = (yearly_total / lag(yearly_total) - 1) )

Year_merchandise <- Year_merchandise %>%
  mutate(value_billions = yearly_total / 1e9)

plot_data <- Year_merchandise %>%
  filter(Category != "Total Trade") 

ggplot() +
  geom_bar(data = plot_data, 
           aes(x = factor(year), y = value_billions, fill = Category),
           stat = "identity", position = position_dodge(width = 0.7), width = 0.6) +
  
  geom_line(data = Year_merchandise %>% filter(Category == "Total Trade"),
            aes(x = factor(year), y = value_billions, group = 1),
            color = "darkred", size = 1.2) +
  geom_point(data = Year_merchandise %>% filter(Category == "Total Trade"),
             aes(x = factor(year), y = value_billions),
             color = "darkred", size = 3) +
  
geom_text(data = plot_data,
            aes(x = factor(year), y = value_billions, label = paste0("$", round(value_billions, 1), "B"), 
                group = Category),
            position = position_dodge(width = 0.7), vjust = -0.5, size = 2.5) +
  
  geom_text(data = Year_merchandise %>% filter(Category == "Total Trade"),
            aes(x = factor(year), y = value_billions, 
            label = paste0("S$ ", round(value_billions, 1), " Bil")),
            vjust = -0.8, color = "darkred", fontface = "bold", size = 2.8) +
  
  geom_text(data = trade_growth %>% filter(!is.na(growth_pct)),
            aes(x = factor(year), y = 0, 
                label = paste0(ifelse(growth_pct > 0, "+", ""), round(growth_pct, 1), "%")),
            vjust = 1.5, color = "gray30", fontface = "italic", size = 2.2) +
  
  scale_fill_manual(values = c("Exports" = "#1f77b4", "Imports" = "#ff7f0e")) +
  
  scale_y_continuous(labels = function(x) paste0("$", x, "B"), 
                     expand = expansion(mult = c(0, 0.15))) +
  
  labs(
    title = "TOTAL MERCHANDISE TRADE 2018 - 2024",
    x = NULL,
    y = "Billions SGD",
    fill = "Category"
  ) +
  
  theme_minimal() +
    theme(
    legend.position = "top",
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold",hjust = 0.5)
 
  )
 

```


1. 
TOTAL MERCHANDISE TRADE AT CURRENT PRICES, 2020 - 2024
pros: 
cons:

sketches:

::: info-box
This analysis utilizes the "Total Merchandise Trade at Current Prices" dataset from the Singapore Department of Statistics (DOS). Data for "Total Merchandise Trade," "Total Merchandise Exports," and "Total Merchandise Imports" were extracted to better analyze recent trends in trade volume and compare fluctuations before and after the COVID-19 pandemic, as well as post-pandemic recovery.
The dataset originally presents time in a Year + Month format. To make it more suitable for annual analysis, the data was first filtered for the years 2018-2024. The mutate() function was used to extract the month and year components, and the summarize() function aggregate the data at the yearly level, calculating the total trade volume for exports and imports.

批判：
The original bar chart employed a container-themed design, aligning with the trade context.   While visually engaging, the color scheme lacked functional utility, hindering multi-dimensional comparisons (e.g., year-over-year trends and import/export differences).   Additionally, the 5-year timeframe limited the ability to contextualize pandemic-related impacts.

我的：
The design combine a line chart and stacked bar charts  This clarifies the relationship between total trade (imports + exports) and individual components. The line chart emphasizes annual trends, while stacked bars breakdown the import and export. A unified color scheme improves cross-year comparisons of imports and exports and increase the space for each year can improve readability.

:::




Import the data
```{r}

servicetrade <- read_xlsx("Trade_Exp_Imp_by_Services_201525.xlsx")
```



```{r}
kable(head(servicetrade))
```



```{r}
servicetrade <- servicetrade %>%
  filter(trimws(`Data Series`) %in% c("Exports Of Services", 
                                      "Manufacturing Services On Physical
                                      Inputs Owned By Others",
                                      "Maintenance And Repair Services",
                                      "Transport", 
                                      "Travel",
                                      "Insurance", 
                                      "Government Goods And Services",
                                      "Construction",
                                      "Financial", "Telecommunications,Computer
                                      & Information",
                                      "Charges For The Use Of Intellectual
                                      Property",
                                      "Personal, Cultural And Recreational",
                                      "Business Management",
                                      "Trade-related", "Other Services")) %>%
  select(`Data Series`, `2024`) %>%
  mutate(value_billions = as.numeric(gsub(",", "", `2024`))/1e3) %>%
  filter(`Data Series` != "Exports Of Services") %>%
  arrange(desc(value_billions))

  ggplot(servicetrade, 
       aes(area = value_billions, 
           fill = value_billions,
           label = paste0(`Data Series`, "\nS$ ", round(value_billions, 1), " Bil"))) +  
  
  geom_treemap(color = "white", size = 1) +
  
  geom_treemap_text(fontface = "bold", colour = "white",
                    place = "centre", grow = TRUE, reflow = TRUE, min.size = 1) +
  
  scale_fill_gradient(low = "#b3e0f7", high = "#052E73") +
  
  labs(title = "2024 Exports of Services by Category",
       fill = "Value (S$ Billion)") +
  
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )


```




好学生样本：

  
```{r}
# Filter the dataset for 2020-2024
Year_merchandise <- Year_merchandise %>%
  filter(year >= 2018 & year <= 2024, 
         grepl("Total Merchandise (Trade|Exports|Imports)", `Data Series`))

# Reshape the data for visualization
Year_merchandise <- Year_merchandise %>%
  mutate(Category = case_when(
    grepl("Total Merchandise Trade", `Data Series`) ~ "Total Trade",
    grepl("Total Merchandise Exports", `Data Series`) ~ "Exports",
    grepl("Total Merchandise Imports", `Data Series`) ~ "Imports"
  )) %>%
  select(year, yearly_total, Category, `Data Series`)

# Calculate year-over-year growth rates for Total Trade
trade_growth <- Year_merchandise %>%
  filter(Category == "Total Trade") %>%
  arrange(year) %>%
  mutate(growth_pct = (yearly_total / lag(yearly_total) - 1) * 100)

# Convert values to billions for easier reading
Year_merchandise <- Year_merchandise %>%
  mutate(value_billions = yearly_total / 1e9)

# Create comparison data - side-by-side bars for imports and exports with trade line
plot_data <- Year_merchandise %>%
  filter(Category != "Total Trade") # Remove total trade for the bars

# Create a more informative and aesthetically pleasing visualization
ggplot() +
  # Side-by-side bars for imports and exports
  geom_bar(data = plot_data, 
           aes(x = factor(year), y = value_billions, fill = Category),
           stat = "identity", position = position_dodge(width = 0.7), width = 0.6) +
  
  # Line for total trade
  geom_line(data = Year_merchandise %>% filter(Category == "Total Trade"),
            aes(x = factor(year), y = value_billions, group = 1),
            color = "darkred", size = 1.2) +
  geom_point(data = Year_merchandise %>% filter(Category == "Total Trade"),
             aes(x = factor(year), y = value_billions),
             color = "darkred", size = 3) +
  
  # Value labels for imports and exports
  geom_text(data = plot_data,
            aes(x = factor(year), y = value_billions, label = paste0("$", round(value_billions, 1), "B"), 
                group = Category),
            position = position_dodge(width = 0.7), vjust = -0.5, size = 3.5) +
  
  # Total trade value labels
  geom_text(data = Year_merchandise %>% filter(Category == "Total Trade"),
            aes(x = factor(year), y = value_billions, 
                label = paste0("Total: $", round(value_billions, 1), "B")),
            vjust = -0.8, color = "darkred", fontface = "bold", size = 3.8) +
  
  # Growth rate annotations
  geom_text(data = trade_growth %>% filter(!is.na(growth_pct)),
            aes(x = factor(year), y = 0, 
                label = paste0(ifelse(growth_pct > 0, "+", ""), round(growth_pct, 1), "%")),
            vjust = 1.5, color = "gray30", fontface = "italic", size = 3.2) +
  
  # Custom color palette for better contrast
  scale_fill_manual(values = c("Exports" = "#1f77b4", "Imports" = "#ff7f0e")) +
  
  # Customize scales and labels
  scale_y_continuous(labels = function(x) paste0("$", x, "B"), 
                     expand = expansion(mult = c(0, 0.15))) +
  
  labs(
    title = "SINGAPORE MERCHANDISE TRADE (2020-2024)",
    subtitle = "Exports, Imports and Total Trade Values",
    caption = "Values in Billions SGD | Percentages show year-over-year growth in Total Trade",
    x = NULL,
    y = "Trade Value (Billions SGD)",
    fill = "Category"
  ) +
  
  # Enhanced theme
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 20)),
    plot.caption = element_text(size = 9, color = "gray30", margin = margin(t = 15)),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 11),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text = element_text(size = 11),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    plot.background = element_rect(fill = "ghostwhite", color = NA),
    panel.background = element_rect(fill = "ghostwhite", color = NA),
    plot.margin = margin(20, 20, 20, 20)
  )
```

