---
title: "Take-Home Exercise 02"
author: "Lin Xinyue"
date: "March 03, 2025"
date-modified: "last-modified"
---
1.Data Preparation 

1.1 Installing R pacakges
```{r}
pacman::p_load(tidyverse,readxl, gifski, gapminder,
               plotly, gganimate, lubridate,
               scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               CGPfunctions, ggHoriPlot, gtsummary,
               easystats,tidyverse, GGally,SmartEDA,
               treemapify,treemap,d3treeR,
               ggExtra,tsibble, feasts,fable, seasonal)
```

1.2 Importing the data

```{r}
merchandisetrade <- read_xlsx("Total_Merchandise_Trade_201525.xlsx")
```



1.3 Examining the data structure

```{r}
kable(head(merchandisetrade))
```


1.4 Data Preparation

For merchandise 

```{r}
month_levels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

Year_merchandise <- merchandisetrade %>%
  pivot_longer(
    cols = matches("\\d{4}\\s[A-Za-z]{3}"), 
    names_to = "date_col",
    values_to = "value"
  ) %>%
  mutate(
    year = as.numeric(str_extract(date_col, "\\d{4}")),
    month = str_extract(date_col, "[A-Za-z]{3}")
  ) %>%
  select(-date_col) %>%
  arrange(year, match(month, month_levels)) %>%
  mutate(month = factor(month, levels = month_levels))

```

```{r}
Year_merchandise <- Year_merchandise %>%
  group_by(`Data Series`, year) %>% 
  summarize(yearly_total = sum(value, na.rm = TRUE), .groups = "drop")
```

```{r}
kable(head(Year_merchandise))
```
```{r}
Year_merchandise %>%
  ExpData(type = 2)
```

visulization

```{r}
Year_merchandise <- Year_merchandise %>%
  filter(year >= 2018 & year <= 2024, 
         grepl("Total Merchandise (Trade|Exports|Imports)", `Data Series`))

Year_merchandise <- Year_merchandise %>%
  mutate(Category = case_when(
    grepl("Total Merchandise Trade", `Data Series`) ~ "Total Trade",
    grepl("Total Merchandise Exports", `Data Series`) ~ "Exports",
    grepl("Total Merchandise Imports", `Data Series`) ~ "Imports"
  )) %>%
  select(year, yearly_total, Category, `Data Series`)

trade_growth <- Year_merchandise %>%
  filter(Category == "Total Trade") %>%
  arrange(year) %>%
  mutate(growth_pct = (yearly_total / lag(yearly_total) - 1) )

Year_merchandise <- Year_merchandise %>%
  mutate(value_billions = yearly_total / 1e9)

plot_data <- Year_merchandise %>%
  filter(Category != "Total Trade") 

ggplot() +
  geom_bar(data = plot_data, 
           aes(x = factor(year), y = value_billions, fill = Category),
           stat = "identity", position = position_dodge(width = 0.7), width = 0.6) +
  
  geom_line(data = Year_merchandise %>% filter(Category == "Total Trade"),
            aes(x = factor(year), y = value_billions, group = 1),
            color = "darkred", size = 1.2) +
  geom_point(data = Year_merchandise %>% filter(Category == "Total Trade"),
             aes(x = factor(year), y = value_billions),
             color = "darkred", size = 3) +
  
geom_text(data = plot_data,
            aes(x = factor(year), y = value_billions, label = paste0("$", round(value_billions, 1), "B"), 
                group = Category),
            position = position_dodge(width = 0.7), vjust = -0.5, size = 2.5) +
  
  geom_text(data = Year_merchandise %>% filter(Category == "Total Trade"),
            aes(x = factor(year), y = value_billions, 
            label = paste0("S$ ", round(value_billions, 1), " Bil")),
            vjust = -0.8, color = "darkred", fontface = "bold", size = 2.8) +
  
  geom_text(data = trade_growth %>% filter(!is.na(growth_pct)),
            aes(x = factor(year), y = 0, 
                label = paste0(ifelse(growth_pct > 0, "+", ""), round(growth_pct, 1), "%")),
            vjust = 1.5, color = "gray30", fontface = "italic", size = 2.2) +
  
  scale_fill_manual(values = c("Exports" = "#1f77b4", "Imports" = "#ff7f0e")) +
  
  scale_y_continuous(labels = function(x) paste0("$", x, "B"), 
                     expand = expansion(mult = c(0, 0.15))) +
  
  labs(
    title = "TOTAL MERCHANDISE TRADE 2018 - 2024",
    x = NULL,
    y = "Billions SGD",
    fill = "Category"
  ) +
  
  theme_minimal() +
    theme(
    legend.position = "top",
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold",hjust = 0.5)
 
  )
 

```


1. 
TOTAL MERCHANDISE TRADE AT CURRENT PRICES, 2020 - 2024

::: info-box
This analysis utilizes the "Total Merchandise Trade at Current Prices" dataset from the Singapore Department of Statistics (DOS). Data for "Total Merchandise Trade," "Total Merchandise Exports," and "Total Merchandise Imports" were extracted to better analyze recent trends in trade volume and compare fluctuations before and after the COVID-19 pandemic, as well as post-pandemic recovery.
The dataset originally presents time in a Year + Month format. To make it more suitable for annual analysis, the data was first filtered for the years 2018-2024. The mutate() function was used to extract the month and year components, and the summarize() function aggregate the data at the yearly level, calculating the total trade volume for exports and imports.

The original bar chart employed a container-themed design, aligning with the trade context.   While visually engaging, the color scheme lacked functional utility, hindering multi-dimensional comparisons (e.g., year-over-year trends and import/export differences).   Additionally, the 5-year timeframe limited the ability to contextualize pandemic-related impacts.

The design combine a line chart and stacked bar charts  This clarifies the relationship between total trade (imports + exports) and individual components. The line chart emphasizes annual trends, while stacked bars breakdown the import and export. A unified color scheme improves cross-year comparisons of imports and exports and increase the space for each year can improve readability.

:::




Import the data
```{r}

servicetrade <- read_xlsx("Trade_Exp_Imp_by_Services_201525.xlsx")
```

```{r}
kable(head(servicetrade))
```


```{r}
servicetrade <- servicetrade %>%
  filter(trimws(`Data Series`) %in% c("Exports Of Services", 
                                      "Manufacturing Services On Physical
                                      Inputs Owned By Others",
                                      "Maintenance And Repair Services",
                                      "Transport", 
                                      "Travel",
                                      "Insurance", 
                                      "Government Goods And Services",
                                      "Construction",
                                      "Financial", "Telecommunications,Computer
                                      & Information",
                                      "Charges For The Use Of Intellectual
                                      Property",
                                      "Personal, Cultural And Recreational",
                                      "Business Management",
                                      "Trade-related", "Other Services")) %>%
  select(`Data Series`, `2024`) %>%
  mutate(value_billions = as.numeric(gsub(",", "", `2024`))/1e3) %>%
  filter(`Data Series` != "Exports Of Services") %>%
  arrange(desc(value_billions))

  ggplot(servicetrade, 
       aes(area = value_billions, 
           fill = value_billions,
           label = paste0(`Data Series`, "\nS$ ", round(value_billions, 1), " Bil"))) +  
  
  geom_treemap(color = "white", size = 1) +
  
  geom_treemap_text(fontface = "bold", colour = "white",
                    place = "centre", grow = TRUE, reflow = TRUE, min.size = 1) +
  
  scale_fill_gradient(low = "#b3e0f7", high = "#052E73") +
  
  labs(title = "2024 Exports of Services by Category",
       fill = "Value (S$ Billion)") +
  
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )


```

::: info-box
The graph uses a tree chart instead of a pie chart. Although the pattern of the previous pie chart was beautiful, different data were classified in different colors and the data part was displayed in larger font, it was difficult to capture useful information at first glance due to the wide variety of types. 
Using the same color tree chart can clearly distinguish which type of service accounts for more in the export proportion, which is more clear and intuitive。
:::







```{r}
DExport_Region <- read_xlsx("Merchandise_Trade_DExport_by Region_201825.xlsx")

Import_Region <- read_xlsx("Merchandise_Trade_Import_by Region.xlsx")

ReExport_Region <- read_xlsx("Merchandise_Trade_ReExport_by Region_201825.xlsx")
```

```{r}

month_levels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

DExport_Region <- DExport_Region %>%
  pivot_longer(
    cols = matches("\\d{4}\\s[A-Za-z]{3}"), 
    names_to = "date_col",
    values_to = "value"
  ) %>%
  mutate(
    year = as.numeric(str_extract(date_col, "\\d{4}")),
    month = str_extract(date_col, "[A-Za-z]{3}")
  ) %>%
  select(-date_col) %>%
  arrange(year, match(month, month_levels)) %>%
  mutate(month = factor(month, levels = month_levels))

```

```{r}
DExport_Region <- DExport_Region %>%
  group_by(`Data Series`, year) %>% 
  summarize(yearly_total = sum(value, na.rm = TRUE), .groups = "drop")
```

Check the data
```{r}
Import_Region %>%
  ExpData(type = 1)
```

Select the data
```{r}
import_long <- Import_Region %>%
  pivot_longer(
    cols = -c(`Data Series`),
    names_to = "Month",
    values_to = "Value"
  ) %>%

  mutate(Month = yearmonth(as.Date(paste0("01 ", Month), format = "%d %Y %b")))


import_ts <- import_long %>%
  as_tsibble(index = Month, key = `Data Series`)

# Visualize the time series data
import_ts %>%
  filter(`Data Series` %in% c("Total All Markets","Asia", "America",
                              "Europe","Oceania", "Africa")) %>%
  ggplot(aes(x = Month, y = Value, color =`Data Series`)) +
  geom_line() +
  labs(title = "Import Data Time Series",
       y = "Value",
       x = "Month") 

# Create individual time series plots for each country
import_ts %>%
  filter(!`Data Series` %in% c("Total All Markets", "America")) %>%
  ggplot(aes(x = Month, y = Value)) +
  geom_line() +
  facet_wrap(~ `Data Series`, scales = "free_y") +
  labs(title = "Import Data by Country",
       y = "Value",
       x = "Month") 
```




Visual Analysis of Seasonality with Seasonal Plot
```{r}
import_ts %>%
  filter(`Data Series` %in% c("Total All Markets","Asia", "America",
                              "Europe","Oceania", "Africa")) %>%
  gg_season(Value) 
```


Visual Analysis of Seasonality with Cycle Plot
```{r}
# Subseries plot
import_ts %>%
  filter(`Data Series` %in% c("Total All Markets","Asia", "America",
                              "Europe","Oceania", "Africa")) %>%
  gg_subseries(Value)
```


Time series decomposition
```{r}

# Define the categories you want to analyze
selected_countries <- c("Total All Markets", "Asia", "America", "Europe", "Oceania", "Africa")

# Compute PACF and plot all categories in one faceted plot
import_ts %>%
  filter(`Data Series` %in% selected_countries) %>%  # Filter all selected categories
  PACF(Value) %>%  # Compute PACF
  autoplot() +  # Facet by 'Data Series'
  labs(title = "PACF Analysis Across Selected Categories")

```

```{r}
import_ts %>%
  filter(`Data Series` == "Total All Markets") %>%
  gg_tsdisplay(Value)
```


Classical Decomposition with feasts
```{r}
import_ts %>%
  filter(`Data Series` == "Total All Markets") %>%
  model(
    classical_decomposition(
      Value, type = "additive")) %>%
  components() %>%
  autoplot()
```

Time Series Data Sampling
```{r}
import_ts <- import_ts %>%
  mutate(Type = if_else(
    Month >= yearmonth("2024 Nov"), 
    "Hold-out",  "Training"))

# Create training data
train_data <- import_ts %>%
  filter(Type == "Training")

# Create test data (hold-out)
test_data <- import_ts %>%
  filter(Type == "Hold-out")


```


```{r}
for (country in c("Total All Markets", "Asia", "America", "Europe", "Oceania", "Africa")) {
  print(paste("STL Decomposition for:", country))
  
  train_data %>%
    filter(`Data Series` == country) %>%
    model(stl = STL(Value)) %>%
    components() %>%
    autoplot() %>%
    print()
}

```

Fitting forecasting models
Fit Simple Exponential Smoothing (SES)
```{r}
fit_ses <- train_data %>%
  filter(`Data Series` %in% c("Total All Markets", "Asia", "America", "Europe",
                              "Oceania", "Africa")) %>%
  model(ses = ETS(Value ~ error("A") + trend("N") + season("N")))

fit_ses

```
Examine Model Assumptions
```{r}
fit_ses %>%
  filter(`Data Series` == "Total All Markets") %>%
  gg_tsresiduals()
```

The model details

```{r}
fit_ses %>%
  report()
```

Fitting ETS Methods with Trend: Holt’s Linear
```{r}
fit_holt <- train_data %>%
  filter(`Data Series` %in% c("Total All Markets", "Asia", "America", "Europe", "Oceania", "Africa")) %>%
  model(`Holt's method` = ETS(Value ~ error("A") + trend("A") + season("N")))
fit_holt %>%
  filter(`Data Series` == "Total All Markets") %>%
  report()
```
check the result
```{r}
fit_holt %>%
  filter(`Data Series` == "Total All Markets") %>%
  gg_tsresiduals()
```
Fitting ETS Methods with Season: Holt-Winters
```{r}
fit_ets_auto <- train_data %>%
  filter(`Data Series` %in% c("Total All Markets", "Asia", "America", "Europe", "Oceania", "Africa")) %>%
  model(ets_auto = ETS(Value))
fit_ets_auto %>% report()
```
Fitting multiple ETS Models

```{r}
fit_ETS <- train_data %>%
  filter(`Data Series` %in% c("Total All Markets", "Asia", "America", "Europe", "Oceania", "Africa")) %>%
  model(
    ses = ETS(Value ~ error("A") + 
                trend("N") + 
                season("N")),
    holt = ETS(Value ~ error("A") + 
                 trend("A") + 
                 season("N")),
    `damped Holt` = 
          ETS(Value ~ error("A") +
                trend("Ad") + 
                season("N")),
        `WH_A` = ETS(
          Value ~ error("A") + 
            trend("A") + 
            season("A")),
        `WH_M` = ETS(Value ~ error("M") 
                         + trend("A") 
                         + season("M"))
    
  )
```


Model Coefficient
```{r}
fit_ETS %>%
  tidy()
```

Model Comparison
```{r}
fit_ETS %>% 
  report()
```

Forecasting future values



```{r}
fit_ETS %>%
  forecast(h = "12 months") %>%
  autoplot(import_ts, 
           level = NULL)
```

Fitting ETS Automatically
```{r}
fit_autoETS <- train_data %>%
  model(ETS(Value))

fit_autoETS %>% report()
```
```{r}
fit_autoETS %>%
  filter(`Data Series` == "Total All Markets") %>% 
  gg_tsresiduals()

```


Forecast the future values
```{r}
fit_autoETS %>%
  filter(`Data Series` == "Total All Markets") %>% 
  forecast(h = "12 months") %>%
  autoplot(train_data)
```
::: info-box
This analyzes merchandise import trends across five regions (America, Asia, Europe, Oceania, and Africa) using time series forecasting methods from the tidyverts method.
The dataset was converted from wide to long format using pivot_longer, enabling time series analysis with tsibble.Missing values and data types were checked using  ExpData to confirm data integrity.
In general,despite fluctuations such as the 2008 financial crisis and COVID-19 in 2019, imports showed a long-term upward trend. Although imports dropped sharply during COVID-19, rebounded strongly post-pandemic.
In terms of different areas, Asia dominated global imports, reflecting its economic scale.Africa had the lowest import values but exhibited rapid growth post-2015.
In terms of the model, SES Produced residuals without significant patterns, indicating adequate fit.
For the future forecast, although the import value is on the rise, it is also difficult to predict the unexpected situation that the market may face. Different forecasting models also apply to different regions

:::

































 
